---
title: "Generating Predictions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{generating-probabilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = F, message = F}
library(herosurv)
```

Any herosurv survival object, or supported third-party distribution object, can be used to generate predicted values

## Survival Probabilities

Predicted survival probabilities can be generated for a given survival distribution and set of times using the `surv_prob` function.

```{r}
  dist1 <- define_surv_param('exp', rate = 0.04)
  surv_prob(dist1, c(0, 1, 2, 3))
```

## Event Probabilities

Event probabilities can be generated for a given survival distributon and set of time intervals using the `event_prob` function.

```{r}
  dist1 <- define_surv_cure('exp', theta = 0.21, rate = 0.04, mixture = TRUE)
  event_prob(dist1, c(0, 1, 2, 3), c(1, 2, 3, 4))
```