---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{herosurv}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = F, message = F}
library(herosurv)
```

The herosurv package makes it easier to perform applied survival modeling in the context of Markov or partitioned survial models. This is accomplished by providing a series functions for defining survival distribution objects as well as a series of S3 methods that can operate on them. These S3 methods also include generics that provide support for survival distributions estimated from supported third-party packages (e.g. survival, flexsurv, flexsurvcure). All survival distribution functions, except those for generating predicted probabilities, return survival distribution objects allowing multiple function calls to be strung together.

## Defining a Simple Survival Distribution

The herosurv package contains a number of functions for defining survival distributions of various types. For example, the `define_surv_param` function allows you to define a parametric survival distribution of a given family with given parameter values.
```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
```

More information on the kinds of survival distributions supported by herosurv can be found in the "Defining Survival Distributions" vignette.

## Printing a Survival Distribution

Printing a herosurv survival distribution will provide a description of its contents.

```{r}
  dist1 <- define_surv_param('exp', rate = 0.013)
  print(dist1)
```

## Generating Survival Probabilities

Survival probabilities can be generated for a distribution using the `surv_prob` function.

```{r}
  dist1 <- define_surv_param('gompertz', shape = 0.5, rate = 0.03)
  surv_prob(dist1, seq(from = 0, to = 20, by = 1))
```

## Plotting

Survival probabilities for a distribution can be graphed using the `plot` method.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  plot(dist1, max_time = 20)
```

## Modifying and Combining Survival Distributions

Survival distributions can be modified using a variety of functions provided by herosurv. For example, the `apply_hr` function can be used to apply a hazard ratio to a survival distribution.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- apply_hr(dist1, 0.45)
  print(dist2)
```

It is also possible to combine multiple survival distributions into one. For example, the `join` function allows two survival distributions to be joined at the specified cutpoint.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- define_surv_param('weibull', shape = 1.4, scale = 4.1)
  dist3 <- join(dist1, 3, dist2) # Join dist2 and dist3 at time 3
  print(dist3)
```

## Composing Survival Distributions

More complicated survival distributions can be created by stringing together multiple herosurv operations. Use of the `%>%` operator from magrittr is supported.

```{r}

  # create an exponential distribution
  dist1 <- define_surv_param('exp', rate = 0.04)
  
  # create a survival distribution based on custom function
  dist2 <- define_surv_func(function(t) pweibull(t, 1.2, 20.1, lower.tail = FALSE))

  # create a Royston & Parmar spline model
  dist3 <- define_surv_spline(
    scale = 'hazard', # spline used to model log cumulative hazards
    -2.08, 2.75, 0.23, # parameters
    -1.62, 0.57, 1.191
  )

  dist4 <- dist1 %>% # take dist1
      apply_hr(0.4) %>% # apply a hazard ratio of 0.4 to it
      join(4, dist2) %>% # join it to dist2 at time 4
      mix(0.25, dist3, 0.75) # mix the result with dist4 with weights of 25% and 75% respectively

  print(dist4)
```