---
title: "herosurv"
output: rmarkdown::html_vignette
description: >
  An introduction to using herosurv. You will learn the basics
  of defining survival distributions, modifying/combining them,
  and using them to generate survival/event probabilities.
vignette: >
  %\VignetteIndexEntry{herosurv}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = F, message = F}
library(herosurv)
```

## Introduction

The herosurv package makes it easier to perform applied survival modeling in the context of Markov or partitioned survial models. This is accomplished by providing a series functions for defining survival distribution objects as well as a series of S3 methods that can operate on them. These S3 methods also include generics that provide support for survival distributions estimated from supported third-party packages (e.g. survival, flexsurv, flexsurvcure). All survival distribution functions, except those for generating predicted probabilities, return survival distribution objects allowing multiple function calls to be strung together.

## Defining Survival Distributions

The herosurv package provides functions for defining a wide variety of survival distributions of different types.

### Defining a Parametric Survival Model

The `define_surv_param` function allows you to define a parametric survival model with specified distribution and parameter values. Parameterization is based on the [flexsurv](https://cran.r-project.org/web/packages/flexsurv/index.html) package.

```{r}
  param_surv <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  print(param_surv)
```

### Defining a Parametric Mixture/Non-Mixture Cure Model

The `define_surv_cure` function allows you to define a parametric mixture/non-mixture cure model with specified distribution, cure fraction, and parameter values.

```{r}
  cure_surv <- define_surv_cure('exp', theta = 0.21, rate = 0.04, mixture = TRUE)
  print(cure_surv)
```

### Defining a Royston & Parmar Spline Model

The `define_surv_spline` function allows you to define a Royston & Parmar restricted cubic spline model with the specified scale, parameter values, and knot times.

```{r}
  spline_dist <- define_surv_spline(
    scale = 'hazard', # spline used to model log cumulative hazards
    -2.08, 2.75, 0.23, # parameters
    -1.62, 0.57, 1.191 # knot times
  )
  print(spline_dist)
```

### Defining a Survival Distribution from Kaplan-Meier Output

The `define_surv_km` function allows you to define a survival distribution based on the times and survival probabilities from a Kaplan-Meier.
```{r}
  km_output <- data.frame(
    times = c(0, 2, 5, 10, 17, 21),
    surv_prob = c(1, 0.98, 0.94, 0.89, 0.78, 0.78)
  )
  km_dist <- define_surv_km(km_output, "times", "surv_prob")
  print(km_dist)
```

### Defining a Survival Distribution from a Life-Table

The `define_surv_lifetable` function allows you to define a survival distribution based on a life-table, starting age, and gender mix.

```{r}
  life_table <- data.frame(
    age_in_years = c(50, 51, 52, 53, 54, 55),
    p_death_male = c(0.0021, 0.0022, 0.0022, 0.0023, 0.0022, 0.0024),
    p_death_female = c(0.0018, 0.0018, 0.0018, 0.0019, 0.0019, 0.0020)
  )
  lifetable_dist <- define_surv_lifetable(
    life_table, # data.frame containing life-table
    51, # starting age
    0.48, # percent male
    age_col = 'age_in_years', # column in life-table containing age
    male_col = 'p_death_male', # column in life-table containing male p(death)
    female_col = 'p_death_female' # column in life-table containing female p(death)
  )
  print(lifetable_dist)
```

### Defining a Survival Distributions from a Custom Function

The `define_surv_func` function allows you to define a survival distribution based on a custom function.

```{r}
  custom_dist <- define_surv_func(function(t) pweibull(t, 1.2, 20.1, lower.tail = FALSE))
  print(custom_dist)
```

### Support for Survival Distributions from Other Packages

In addition to the functions provided by herosurv, you can also create survial distributions recognized by herosurv using supported functions from other packages.

```{r}
  library(survival)
  library(flexsurv)
  survfit_dist <- survfit(Surv(recyrs, censrec)~1, data = bc)
  flexsurv_dist <- flexsurvreg(Surv(recyrs, censrec)~1, data = bc, dist = 'lnorm')
```

## Modifying & Combining Survival Distributions

Once you've created survival distributions, herosurv provides various methods for combining modifying them.

### Applying a Hazard Ratio:

Hazard ratios can be applied to a survival distribution using the `apply_hr` function.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- apply_hr(dist1, 0.85)
  print(dist2)
```

### Applying an Odds Ratio

Odds ratios can be applied to a survival distribution using the `apply_or` function.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- apply_or(dist1, 0.45)
  print(dist2)
```

### Applying an Accleration Factor

Accleration factors, which proportionally scale event times, can be applied using the `apply_af` function.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- apply_af(dist1, 0.75)
  print(dist2)
```

### Applying a Shift in Time

Hazards for a distribution can be shifted forward or backward by a fixed amount of time using the `apply_shift` function. Note that a positive value of shift will move hazards forward and a negative value will shift hazards backward (resulting in an initial period with no event risk).

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- apply_shift(dist1, 2.1)
  print(dist2)
```


### Mixing Distributions Accordingy to Specified Weights

Survival distributions can be mixed according to specified weights using the `mix` function.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- define_surv_param('weibull', shape = 1.4, scale = 4.1)
  dist3 <- mix(dist1, 0.8, dist2, 0.2) # Mix of 80% dist1 and 20% dist2
  print(dist3)
```

### Joining Two or More Distributions at Specified Cutpoints

Survival distributions can be joined together at specified cutpoints using the `join` function.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- define_surv_param('weibull', shape = 1.4, scale = 4.1)
  dist3 <- join(dist1, 3, dist2) # Join dist2 and dist3 at time 3
  print(dist3)
```

### Combining Two or More Distributions as Indepedent Risks

Survival distributions can be combined as independent risks using the `add_hazards` function. This results in a survival distribution with survival function equal to the product of the survival functions of combined distributions.

```{r}
  dist1 <- define_surv_param('weibull', shape = 1.2, scale = 5.3)
  dist2 <- define_surv_param('weibull', shape = 1.4, scale = 4.1)
  dist3 <- add_hazards(dist1, dist2) # combined hazards of dist1 and dist2
  print(dist3)
```

### Setting Covariates For Model with Predictors

Models fitted using `survfit` or `flexsurvreg` with predictors can be used to generate a survival distribution representing a subject or set of subjects using `set_covariates`.

```{r}
library(flexsurv)
library(tibble)

# Fit a model using bc dataset using good as predictor
print(as_tibble(bc))
fs1 <- flexsurvreg(Surv(recyrs, censrec)~group, data = bc, dist = 'weibull')

# Survival distribution representing patient w/ group = "Good"
good_model <- set_covariates(fs1, data.frame(group = 'Good'))
print(good_model)

# Survival distribution with patient representing cohort w/
# equal mix of group = "Good" and group = "Poor"
good_and_poor_model <- set_covariates(fs1, data.frame(group = c('Good', 'Poor')))
print(good_and_poor_model)

```

## Survival Distribution Composition

More complicated survival distributions can be created by stringing together multiple herosurv operations. Use of the `%>%` operator from magrittr is supported.

```{r}
  dist1 <- define_surv_param('exp', rate = 0.04)
  dist2 <- define_surv_func(function(t) pweibull(t, 1.2, 20.1, lower.tail = FALSE))
  dist3 <- define_surv_spline(
    scale = 'hazard', # spline used to model log cumulative hazards
    -2.08, 2.75, 0.23, # parameters
    -1.62, 0.57, 1.191
  )

  dist4 <- dist1 %>% # take dist1
      apply_hr(0.4) %>% # apply a hazard ratio of 0.4 to it
      join(4, dist2) %>% # join it to dist2 at time 4
      mix(0.25, dist3, 0.75) # mix the result with dist4 with weights of 25% and 75% respectively
  print(dist4)
```

## Generating Predicted Values

### Survival Probabilities

Predicted survival probabilities can be generated for a given survival distribution and set of times using the `surv_prob` function.

```{r}
  dist1 <- define_surv_param('exp', rate = 0.04)
  surv_prob(dist1, c(0, 1, 2, 3))
```

### Event Probabilities

Event probabilities can be generated for a given survival distributon and set of time intervals using the `event_prob` function.

```{r}
  dist1 <- define_surv_cure('exp', theta = 0.21, rate = 0.04, mixture = TRUE)
  event_prob(dist1, c(0, 1, 2, 3), c(1, 2, 3, 4))
```